On The Impact of Atoms of Confusion in JavaScript Code

### 2021-10-22 4_Result.tex
Made a pass through Section 4.

Our first study investigates the impact of atoms of confusion while developers try to understand JavaScript code.
We estimate this impact considering two perspectives: \emph{misunderstanding rate} (number of wrong answers)
and \emph{time} (how long to provide a correct answer).
We received full answers to our survey from 140 participants (a total of 70 replicas).
All participants had taken at least some university course or hold a bachelor
degree or equivalent, meaning the participants have had some level of formal education in programming. Twenty four participants hold either a master's degree (21 participants) or a doctorate degree (three participants). 
Considering the programming experience of our respondents, 19\% have more than ten years of programming experience, 37\% have between four and ten years of experience, 37\% have between one and four years of experience, and 7\% have less than one year of programming experience.  
Accordingly, we characterize the effect of atoms of confusion in JavaScript
code taking into account the perceptions of both novice and experienced developers.

@@

Considering Table~\ref{tab:difference-correctness}, the clean versions of six code snippets present at least a 15\% improvement in answer correctness when compared with the confusing versions. In particular, the presence of the \emph{Comma Operator} atom presents the highest impact on misunderstanding. 
Frequently used constructs and idioms, such as \emph{Post Increment} and \emph{Omitted Curly Braces} (see Section~\ref{sec:msr-results}), also result in many mistakes. %introduce high degrees of confusion.
The boxplot of Figure~\ref{fig:boxplotcorrectness} shows a non-negligible decrease in the average number of incorrect answers when observing the clean versions of the code snippets. Also, the sample of responses with no atoms had almost no dispersion, which supports the
argument that the non-confusing versions of the code snippets are easier to evaluate correctly.

@@

\begin{table}[htbp]
\caption{Summary of the correctness analysis}
\label{tab:difference-correctness}
\centering{
\begin{scriptsize}
\begin{tabular}{lccc} \toprule
  Atom & Confusing Code & Clean Code & $\Delta$(\%)  \\ \midrule
  Comma Operator                &  28 &  65 & +132   \\ 
  Automatic Semicolon Insertion &  32 &  68 & +112   \\ 
  Post-Increment                &  48 &  64 & +33    \\ 
  Omitted Curly Braces          &  47 &  58 &  +23   \\
  Assignment as Value           &  56 &  68 &  +21   \\ 
  Implicit Predicate            &  58 &  68 &  +17   \\ 
  Logic as Control Flow         &  41 &  48 &  +17   \\ 
  Conditional Operator              &  60 &  66 &  +10   \\ 
  Pre-Increment                 &  50 &  53 &  +6    \\ 
  Arithmetic as Logic           &  64 &  63 &  -2    \\ \bottomrule

@@

\begin{figure}[b!]
\noindent
\vspace*{-1.4cm}
 \centering
 \includegraphics[width=\columnwidth]{images/wrong-answers-plot-1.pdf}
 \vspace*{-1.3cm}
 \caption{Number of wrong answers of each subject.}
 \label{fig:boxplotcorrectness}
 \end{figure}

@@

\subsubsection*{Hypothesis Testing and Effect Size}

@@

Finally, we also use the \emph{Binomial Generalized Logistic Regression} analysis to investigate if either \emph{participant education} or \emph{participant experience} impact correctness. The findings suggest that \emph{participant experience} impacts on the results related to correctness (see Figure~\ref{fig:correctness-over-experience}). Even though the presence of atoms of confusion reduces the number of correct answers to all \emph{experience groups}, this effect is not uniform. For instance, novice developers provide a higher number of wrong answers for the confusing version of the code (57.8\%), while developers with more than four years of experience provide more than 70\% of correct answers even when evaluating the confusing version of the code snippets. It is also important to note that, independently of the \emph{experience group}, developers tend to provide more than 84\% of correct answers while predicting the outputs for the clean versions of the code. We replicate the \emph{Pearson's Chi-squared test} for each experience group and found that the statistical difference is less significant for those developers with more than ten years of experience. Differently from \emph{participant experience}, the factor \emph{participant education} does not change substantially the correctness of the answers.

\begin{figure}
  \includegraphics[scale=0.3]{images/correctness-by-experience}
  \caption{Impact of atoms of confusion on the correctness (over the participant experience groups)}
  \label{fig:correctness-over-experience}
\end{figure}

@@

Comparing to previous research, three 
atoms of confusion that Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} confirmed
for \clang and \cpplang do not lead to a statistically significant impact
on correctness in our study: Logic as Control Flow,
Conditional Operator, and Pre-Increment. Our results
support the Gopstein et al. findings for the remaining
atoms. Considering the work of Langhout and Aniche~\cite{Langhout:2021:ACJ},
their work investigates six atom candidates for Java that
we also explore here for JavaScript. Among these atom
candidates, we achieve similar conclusions with their
study for four atom candidates: Post Increment,
Conditional Operator, Arithmetic as Logic, and
Omitted Curly Braces---both studies confirmed only the former
as an atom of confusion. Our findings diverged from
the work of Langhout and Aniche for two atoms: Pre Increment and
Logic as Control Flow.

@@

\begin{table}[htbp]
\caption{Time in seconds to submit a correct answer}
\centering{
\begin{scriptsize}
\begin{tabular}{lccc} \toprule
Atom & Confusing Code & Clean Code & $\Delta$(\%) \\ \midrule
Comma Operator & 87.67 & 20.84 & -76.23 \\ 
Automatic Semicolon Insertion & 46.08 & 22.04 & -52.17 \\ 
Post-Increment & 28.70 & 25.67 & -10.56 \\ 
Omitted Curly Braces & 48.85 & 30.00 & -38.58 \\ 
Assignment as Value & 52.47 & 48.95 & -6.71 \\ 
Implicit Predicate & 36.24 & 24.01 & -33.75 \\ 
Logic as Control Flow & 108.94 & 51.07 & -53.12 \\ 
Conditional Operator & 41.80 & 39.34 & -5.90 \\ 
Pre-Increment & 30.71 & 42.45 & 38.19 \\  
Arithmetic as Logic & 28.82 & 37.20 & 29.06 \\ \bottomrule
\end{tabular}
\label{tab:difference-time-taken}
\end{scriptsize}
}
\end{table}

 % latex table generated in R 4.0.4 by xtable 1.8-4 package
 % Thu Apr 22 07:58:21 2021
 \begin{table*}[ht]
\caption{Hypotheses Testing (misunderstanding and time)}

 \centering
 \begin{tabular}{lccc|cc}
   \toprule
Atom  & Chi-square test & Odds Ratio          & CI & Mann-Whitey U test & Cliff's Delta \\ 
     &         & (misunderstanding)  &                    & (workload)             & \\ \midrule
Comma Operator & 1.1727e-10 & 19.02 & (6.60, 68.22) & 5.8249e-08 & -0.53 \\ 
Automatic Semicolon Insertion & 5.8352e-11 & 39.33 & (9.21, 356.62) & 0.09802 & -0.16 \\ 
Post-Increment & 0.0015279 & 4.83 & (1.73, 15.74) & 0.12006 & 0.15 \\ 
Omitted Curly Braces & 0.050962 & 2.35 & (1.00, 5.77) & 0.92529 & -0.01 \\ 
Assignment as Value & 0.0034775 & 8.39 & (1.81, 79.12) & 0.82681 & 0.02 \\ 
Implicit Predicate & 0.01123 & 6.95 & (1.46, 66.56) & 0.0029039 & -0.29 \\ 
Logic as Control Flow & 0.292 & 1.54 & (0.73, 3.28) & 7.5862e-05 & -0.39 \\ 
Conditional Operator & 0.15896 & 2.73 & (0.74, 12.57) & 0.24073 & -0.12 \\ 
Pre-Increment & 0.70147 & 1.25 & (0.55, 2.85) & 0.0062981 & 0.27 \\ 
Arithmetic as Logic & 1 & 0.84 & (0.22, 3.12) & 0.01723 & 0.23 \\ 
    \bottomrule
 \end{tabular}
 \label{tab:hypothesis-testing}
 \end{table*}

\subsubsection*{Hypothesis Testing}
We use the non-parametric \emph{Mann-Whitney U test}, also known as Wilcoxon rank-sum test, to
investigate the null hypothesis that developers 
spend the same amount of time to correctly
predict the outcome of a code snippet, regardless
of evaluating the confusing or clean version
of the code. The results of Table~\ref{tab:difference-time-taken}
suggests that we should refute the null hypotheses for five atom candidates (Comma Operator, Implicit Predicate, Logic as Control Flow, Pre-Increment, and Arithmetic as Logic), after applying the Benjamini-Hochberg correction with a false discovery rate of 5\%. For the first three, the analysis suggests that the participants need more time to predict the outcome of the code snippet in the confusing code version. For the atom candidates Arithmetic as Logic and Pre-Increment, the participants take less time to predict the outcome of the code snippets in the confusing version. We also computed the effect size using Cliff's Delta statistic. In the table, negative effect sizes suggest that it took less time to correctly predict the output of clean versions of the snippets.  Considering the thresholds established by Romano et al.~\cite{Romano:2006:ASO} (0.147 $<$ $|$d$|$ $<$ 0.33 small, $|$d$|$ $<$ 0.474 medium, otherwise large), we found a large effect for the Comma Operator atom candidate; a medium effect size for Logic as Control Flow, and small effect sizes for Automatic Semicolon Insertion, Implicit Predicate, Arithmetic as Logic, Post-, and Pre-Increment.

@@

Next, we summarize the results of our second study (interview with
practitioners), which
we use to contrast the findings discussed here with the opinion
of software developers about their preferences regarding the confusing
or clean versions of the code snippets we used in our experiment.

@@

@@

\subsubsection*{The Participants' perceptions of the atom candidates}
Supporting the results of the survey, Table~\ref{tab:interview-results1}
 shows that for eight out of the
10 scenarios surveyed, the majority of the respondents prefer the version of the
code without the atom candidate.
In no case the \emph{neutral}
ratio was higher than the option for the clean version.
%% An example entry is contained in Appendix~\ref{}. 
%% For a full listing of the code snippets, visit the paper repository. 
Only for the Conditional Operator atom candidate the participants prefer the Confusing version, instead of the clean one. We show the code snippets for the Conditional Operator in Figure~\ref{code:ternary}. Even for this atom candidate, some participants who preferred the left-hand side version (confusing) still believed that the right-hand side version (clean) was more readable. The following quotes were extracted from the transcripts with
% three interviewees:
two interviewees:

@@

\noindent\begin{minipage}{.45\textwidth}
\begin{lstlisting}[language=JavaScript, caption=\emph{Left-hand side} (using the \emph{Conditional Operator} atom).]
let config = {size: 3, isActive: false};
const_config = config.isActive === true 
             ? config 
             : {size: 10};
console.log(_config.size);
\end{lstlisting}
\end{minipage}\hfill
\begin{minipage}{.45\textwidth}
\begin{lstlisting}[language=JavaScript, caption=\emph{Right-hand side} (without the atom).]
let config = {size: 3, isActive: false}
let _config;
if(config.isActive === true) {
  _config = config;
}
else{
   _config = {size: 10};
}
console.log(_config.size);
\end{lstlisting}
\end{minipage}
\caption{Example of a pair of code snippets used in the interview. This case explores the use of the \emph{Conditional Operator atom of confusion}}
\label{code:ternary}
\vspace{-0.2cm}
\end{figure*}

@@

\begin{mq}
\emph{``When I am programming, I write code with the conditional operator, [...], but, to be honest, I still think that the 
[the code using the \rhs version] 
is easier to understand''}.
\end{mq}

@@

The Pre-Increment atom candidate also caused such
divide in one of the interviewees, who
regarded the clean version as simpler to understand, but would still opted to write
code with the atom. In contrast, %to such opinion,
one of the participants found the version with the atom candidate more elegant, but recognized it was less readable, and was willing to sacrifice elegance for readability.

@@

Two other developers implicitly touched upon the callback pattern, mentioning that it can be difficult to understand \emph{asynchronous} programming in JavaScript.
%Since what is really happening at a lower level is that the JavaScript engine creates a callback stack that is separate from the main execution stack, and that callback functions are only executed when the main execution stack is empty, the concepts of asynchronous events and callbacks are inseparable in the language, and any abstractions for callback functions, such as promises and async/await syntax only hide the pattern.
We also found other idioms that are JavaScript atom candidates, including
\emph{property access} via array subscription (\lstinline[language=javascript]{V1['P1']} instead of \lstinline[language=javascript]{V1.P1}) and \emph{arrow functions} (see listings~\ref{arr1} and~\ref{arr2}).  As future work, we aim at investigating whether or not these atom candidates are more likely to introduce confusion in JavaScript code. The following section presents the results of our mining software repository effort, whose goal is to reveal how frequent atom candidates are in open source code.

@@

We mined \minedprojects open source JavaScript repositories to understand how
often atoms of confusion arise in real software. 
Similarly to previous studies~\cite{DBLP:conf/msr/GopsteinZFC18,DBLP:journals/ese/MedeirosLAAKRG19}, which investigate
the prevalence of atoms of confusion in open source \clang and \cpplang projects, we found that atoms of confusion
frequently arise in JavaScript open source systems. For instance, the six most frequently found atoms
occur in at least 80\% of the projects. Considering the extremes, atom candidates Conditional Operator and Implicit Predicate were found in all repositories we mined,
while Comma Operator occurred in only 15\% of them (as seen in Table~\ref{tab:occurrences-summary}).

@@

% latex table generated in R 4.0.3 by xtable 1.8-4 package
% Tue Apr 13 09:34:35 2021
\begin{table}[ht]
\centering
\caption{Summary of atom candidate occurrences on our dataset.}
\setlength\tabcolsep{2pt} % default value: 6pt
\begin{tabular}{lrc}%c}
  \toprule
Atom & Projects & Occurr./KLOC \\%& Frequency \\ 
  \midrule
Implicit Predicate & 100\% & 19.89 \\%&  intensively used \\ 
  Conditional Operator & 100\% & 10.16 \\%&  intensively used \\ 
  Omitted Curly Braces & 91.67\% & 6.61 \\%&  intensively used \\ 
  Post Increment & 90.28\% & 5.62 \\%&  intensively used \\ 
  Pre Increment & 83.33\% & 1.04 \\%& commonly used \\ 
  Assignment as Value & 81.94\% & 1.02 \\%& commonly used \\ 
  Logic as Control Flow & 56.94\% & 0.95 \\%& commonly used \\ 
  Automatic Semicolon Insertion & 33.33\% & 0.13 \\%& commonly used \\ 
  Arithmetic as Logic & 23.61\% & 0.02 \\%& rarely used \\ 
  Comma Operator & 15.28\% & 0.03 \\%& rarely used \\
   \bottomrule
\end{tabular}
\label{tab:occurrences-summary}
\end{table}

@@

The results of our survey and interviews suggest that the Conditional Operator does not contribute significantly as a source of misunderstanding (increasing the number of wrong answers in 9\% of the cases, according to our survey). Nonetheless, the Post-Increment Expression and Automatic Semicolon Insertion atoms are listed in the top four sources of misunderstanding  (Table~\ref{tab:difference-correctness}), and also frequently appear in open source systems. As such, fixing the the Post-Increment and Automatic Semicolon Insertion might improve the readabiliy of large amounts of code in JavaScript projects.

>>>>

Our first study investigates the impact of atoms of confusion while developers try to understand JavaScript code.
We estimate this impact considering two perspectives: \emph{misunderstanding rate} (number of wrong answers)
and \emph{time} (how long to provide a correct answer).
We received full answers to our survey from 140 participants (a total of 70 replicas).
All participants had taken at least some university course or hold a bachelor
degree or equivalent, meaning the participants have had some level of formal education in programming. In addition, 21 participants hold a master's degree and three a doctorate degree. 
Considering the programming experience of our respondents, 19\% have more than ten years of programming experience, 37\% have between four and ten years of experience, 37\% have between one and four years of experience, and 7\% have less than one year of programming experience.  
Accordingly, we characterize the effect of atoms of confusion in JavaScript
code taking into account the perceptions of both novice and experienced developers.

@@

Considering Table~\ref{tab:difference-correctness}, the clean versions of six code snippets present at least a 15\% improvement in answer correctness when compared with the confusing versions. In particular, the presence of the \emph{Comma Operator} atom exhibits the highest impact on misunderstanding. 
Frequently used constructs and idioms, such as \emph{Post Increment} and \emph{Omitted Curly Braces} (see Section~\ref{sec:msr-results}), also result in many mistakes. %introduce high degrees of confusion.
The boxplot of Figure~\ref{fig:boxplotcorrectness} shows a non-negligible decrease in the average number of incorrect answers when observing the clean versions of the code snippets. Also, the sample of responses with no atoms had almost no dispersion, which supports the argument that the non-confusing versions of the code snippets are easier to evaluate correctly.

@@

\begin{table}[htbp]
\caption{Summary of the correctness analysis}
\label{tab:difference-correctness}
\centering{
\begin{scriptsize}
\begin{tabular}{lccc} \toprule
  Atom & Confusing & Clean & $\Delta$(\%)  \\ \midrule
  Comma Operator                &  28 &  65 & +132   \\ 
  Automatic Semicolon Insertion &  32 &  68 & +112   \\ 
  Post-Increment                &  48 &  64 & +33    \\ 
  Omitted Curly Braces          &  47 &  58 &  +23   \\
  Assignment as Value           &  56 &  68 &  +21   \\ 
  Implicit Predicate            &  58 &  68 &  +17   \\ 
  Logic as Control Flow         &  41 &  48 &  +17   \\ 
  Conditional Operator              &  60 &  66 &  +10   \\ 
  Pre-Increment                 &  50 &  53 &  +6    \\ 
  Arithmetic as Logic           &  64 &  63 &  -2    \\ \bottomrule

@@

\begin{figure}[b!]
\noindent
\vspace*{-1.4cm}
 \centering
 \includegraphics[width=\columnwidth]{images/wrong-answers-plot-1.pdf}
 \vspace*{-1.3cm}
 \caption{Number of wrong answers of each subject.}
 \label{fig:boxplotcorrectness}
 \end{figure}\castor{Fiz de tudo para ajeitar a figura porque o texto embaixo deveria ser "Clean versions" e "Confusing versions" mas não sei porque não funcionou. Até mudei a figura. Não sei o que está havendo.}

@@

\subsubsection*{Statistical analysis}

@@

Finally, we also use a \emph{Binomial Generalized Logistic Regression} analysis to investigate if either \emph{participant education} or \emph{participant experience} impact correctness. The findings suggest that \emph{participant experience} impacts on the results related to correctness (Figure~\ref{fig:correctness-over-experience}). Even though the presence of atoms of confusion reduces the number of correct answers to all \emph{experience groups}, this effect is not uniform. For instance, novice developers (up to four years of programming experience) provide a higher number of wrong answers for the confusing version of the code (57.8\%), while developers with more than four years of experience provide more than 70\% of correct answers even when evaluating the confusing version of the code snippets. It is also important to note that, independently of the \emph{experience group}, developers tend to provide more than 84\% of correct answers while predicting the outputs for the clean versions of the code. We replicate the \emph{Pearson's Chi-squared test} for each experience group and find that the statistical difference is less significant for those developers with more than ten years of experience. Differently from \emph{participant experience}, the factor \emph{participant education} does not significantly change the correctness of the answers.\castor{What does "for each experience group" mean above?}

\begin{figure}
  \includegraphics[scale=0.3]{images/correctness-by-experience}
  \caption{Impact of atoms of confusion on the correctness (over the participant experience groups)}
  \label{fig:correctness-over-experience}
\end{figure}
\castor{This figure is missing. Is that expected?}

@@

Comparing to previous research, three 
atoms of confusion that Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} confirmed
for \clang and \cpplang do not lead to a statistically significant impact
on correctness in our study: Logic as Control Flow,
Conditional Operator, and Pre-Increment. Our results
support the Gopstein et al. findings for the remaining
atoms. The work of Langhout and Aniche~\cite{Langhout:2021:ACJ} investigates six atom candidates for Java that we also explore here for JavaScript. Among these atom
candidates, we achieve similar conclusions to their
study for four atom candidates: Post Increment,
Conditional Operator, Arithmetic as Logic, and
Omitted Curly Braces---both studies confirmed only the former
as an atom of confusion. Our findings diverged from
the work of Langhout and Aniche for two atoms: Pre Increment and
Logic as Control Flow.

@@

\begin{table}[htbp]
\caption{Time in seconds to submit a correct answer}
\centering{
\begin{scriptsize}
\begin{tabular}{lccc} \toprule
Atom & Confusing  & Clean & $\Delta$(\%) \\ \midrule
Comma Operator & 87.67 & 20.84 & -76.23 \\ 
Automatic Semicolon Insertion & 46.08 & 22.04 & -52.17 \\ 
Post-Increment & 28.70 & 25.67 & -10.56 \\ 
Omitted Curly Braces & 48.85 & 30.00 & -38.58 \\ 
Assignment as Value & 52.47 & 48.95 & -6.71 \\ 
Implicit Predicate & 36.24 & 24.01 & -33.75 \\ 
Logic as Control Flow & 108.94 & 51.07 & -53.12 \\ 
Conditional Operator & 41.80 & 39.34 & -5.90 \\ 
Pre-Increment & 30.71 & 42.45 & 38.19 \\  
Arithmetic as Logic & 28.82 & 37.20 & 29.06 \\ \bottomrule
\end{tabular}
\label{tab:difference-time-taken}
\end{scriptsize}
}
\end{table}

 % latex table generated in R 4.0.4 by xtable 1.8-4 package
 % Thu Apr 22 07:58:21 2021
 \begin{table*}[ht]
\caption{Hypotheses Testing (misunderstanding and time). Asterisks ($^{*}$) indicate a statistically significant difference.}

 \centering
 \begin{tabular}{lccc|cc}
   \toprule
Atom  & Chi-square test & Odds Ratio          & CI & Mann-Whitey U test & Cliff's Delta \\ 
     &         & (misunderstanding)  &                    & (workload)             & \\ \midrule
Comma Operator & 1.1727e-10$^{*}$ & 19.02 & (6.60, 68.22) & 5.8249e-08$^{*}$ & -0.53 \\ 
Automatic Semicolon Insertion & 5.8352e-11$^{*}$ & 39.33 & (9.21, 356.62) & 0.09802 & -0.16 \\ 
Post-Increment & 0.0015279$^{*}$ & 4.83 & (1.73, 15.74) & 0.12006 & 0.15 \\ 
Omitted Curly Braces & 0.050962 & 2.35 & (1.00, 5.77) & 0.92529 & -0.01 \\ 
Assignment as Value & 0.0034775$^{*}$ & 8.39 & (1.81, 79.12) & 0.82681 & 0.02 \\ 
Implicit Predicate & 0.01123$^{*}$ & 6.95 & (1.46, 66.56) & 0.0029039$^{*}$ & -0.29 \\ 
Logic as Control Flow & 0.292 & 1.54 & (0.73, 3.28) & 7.5862e-05$^{*}$ & -0.39 \\ 
Conditional Operator & 0.15896 & 2.73 & (0.74, 12.57) & 0.24073 & -0.12 \\ 
Pre-Increment & 0.70147 & 1.25 & (0.55, 2.85) & 0.0062981$^{*}$ & 0.27 \\ 
Arithmetic as Logic & 1 & 0.84 & (0.22, 3.12) & 0.01723$^{*}$ & 0.23 \\ 
    \bottomrule
 \end{tabular}
 \label{tab:hypothesis-testing}
 \end{table*}

\subsubsection*{Statistical analysis}
We use the non-parametric \emph{Mann-Whitney U test}, also known as Wilcoxon rank-sum test, to
investigate the null hypothesis that developers 
spend the same amount of time to correctly
predict the outcome of clean and confusing versions of a code snippet.
%, regardless of evaluating the confusing or clean version of the code.
The results of Table~\ref{tab:difference-time-taken}
suggest that we should refute the null hypotheses for five atom candidates (Comma Operator, Implicit Predicate, Logic as Control Flow, Pre-Increment, and Arithmetic as Logic), after applying the Benjamini-Hochberg correction with a false discovery rate of 5\%. For the first three, the analysis suggests that the participants need more time to predict the outcome of the code snippet in the confusing code version. For the atom candidates Arithmetic as Logic and Pre-Increment, the participants take less time to predict the outcome of the code snippets in the confusing version. We also computed the effect size using Cliff's Delta statistic. In the table, negative effect sizes suggest that it took less time to correctly predict the output of clean versions of the snippets.  Considering the thresholds established by Romano et al.~\cite{Romano:2006:ASO} (0.147 $<$ $|$d$|$ $<$ 0.33 small, $|$d$|$ $<$ 0.474 medium, otherwise large), we found a large effect for the Comma Operator atom candidate; a medium effect size for Logic as Control Flow, and small effect sizes for Automatic Semicolon Insertion, Implicit Predicate, Arithmetic as Logic, Post-, and Pre-Increment.

@@

@@

In this section we present the results of the interviews with the practitioners. 
We contrast the findings of the survey presented in Section~\ref{sec:survey-resuts} here with the opinion
of software developers about their preferences regarding the confusing
or clean versions of the code snippets we used in our experiment.

@@

\subsubsection*{The Participants' perceptions of the atom candidates} Supporting the results of the survey, Table~\ref{tab:interview-results1}
 shows that for eight out of the
10 scenarios surveyed, the majority of the respondents prefer the version of the
code without the atom candidate.
In no case the \emph{neutral}
ratio was higher than the option for the clean version.
%% An example entry is contained in Appendix~\ref{}. 
%% For a full listing of the code snippets, visit the paper repository. 
Only for the Conditional Operator atom candidate the participants prefer the confusing version, instead of the clean one. Figure~\ref{code:ternary} shows the code snippets for the corresponding confusing and clean versions. For this atom candidate, some participants who preferred the left-hand side version (confusing) still believed that the right-hand side version (clean) was more readable. The following quotes were extracted from the transcripts with
% three interviewees:
two interviewees:

@@

\noindent\begin{minipage}{.45\textwidth}
\begin{lstlisting}[language=JavaScript, caption=\emph{Left-hand side} (using the \emph{Conditional Operator} atom).]
let config = {size: 3, isActive: false};
const_config = config.isActive === true 
             ? config 
             : {size: 10};
console.log(_config.size);
\end{lstlisting}
\end{minipage}\hfill
\begin{minipage}{.45\textwidth}
\begin{lstlisting}[language=JavaScript, caption=\emph{Right-hand side} (without the atom).]
let config = {size: 3, isActive: false}
let _config;
if(config.isActive === true) {
  _config = config;
}
else{
   _config = {size: 10};
}
console.log(_config.size);
\end{lstlisting}
\end{minipage}
\caption{Example of a pair of code snippets used in the interview pertaining to the \emph{Conditional Operator} atom candidate}
\label{code:ternary}
\vspace{-0.2cm}
\end{figure*}

@@

\begin{mq}
\emph{``When I am programming, I write code with the conditional operator, [...], but, to be honest, I still think that 
[the code using the \rhs version] 
is easier to understand''}.
\end{mq}

@@

The Pre-Increment atom candidate also caused such
divide in one of the interviewees, who
regarded the clean version as simpler to understand, but would still opt to write
code with the atom. In contrast, %to such opinion,
one of the participants found the version with the atom candidate more elegant, but recognized it was less readable, and was willing to sacrifice elegance for readability.

@@

Two other developers implicitly touched upon the callback pattern, mentioning that it can be difficult to understand \emph{asynchronous} programming in JavaScript.
%Since what is really happening at a lower level is that the JavaScript engine creates a callback stack that is separate from the main execution stack, and that callback functions are only executed when the main execution stack is empty, the concepts of asynchronous events and callbacks are inseparable in the language, and any abstractions for callback functions, such as promises and async/await syntax only hide the pattern.
We also found other idioms that are JavaScript atom candidates, including
\emph{property access} via array subscription (\lstinline[language=javascript]{V1['P1']} instead of \lstinline[language=javascript]{V1.P1}) and \emph{arrow functions} (see listings~\ref{arr1} and~\ref{arr2}).  As future work, we aim at investigating whether or not these atom candidates are more likely to introduce confusion in JavaScript code.

@@

In this section we present the results of our mining software repository effort, whose goal is to reveal how prevalent atom candidates are in open source JavaScript projects. 
We mined \minedprojects open source JavaScript repositories to understand how
often atoms of confusion arise in real software. 
Similarly to previous studies~\cite{DBLP:conf/msr/GopsteinZFC18,DBLP:journals/ese/MedeirosLAAKRG19}, which investigate
the prevalence of atoms of confusion in open source \clang and \cpplang projects, we found that atoms of confusion
frequently arise in JavaScript open source systems. For instance, the six most frequently found atoms
occur in at least 80\% of the projects. Considering the extremes, atom candidates Conditional Operator and Implicit Predicate were found in all repositories we mined,
while Comma Operator occurred in only 15\% of them (as seen in Table~\ref{tab:occurrences-summary}).

@@

% latex table generated in R 4.0.3 by xtable 1.8-4 package
% Tue Apr 13 09:34:35 2021
\begin{table}[ht]
\centering
\caption{Summary of atom candidate occurrences in our dataset.}
\setlength\tabcolsep{2pt} % default value: 6pt
\begin{tabular}{lrc}%c}
  \toprule
Atom & Projects & Occurr./KLOC \\%& Frequency \\ 
  \midrule
Implicit Predicate & 100\% & 19.89 \\%&  intensively used \\ 
  Conditional Operator & 100\% & 10.16 \\%&  intensively used \\ 
  Omitted Curly Braces & 91.67\% & 6.61 \\%&  intensively used \\ 
  Post Increment & 90.28\% & 5.62 \\%&  intensively used \\ 
  Pre Increment & 83.33\% & 1.04 \\%& commonly used \\ 
  Assignment as Value & 81.94\% & 1.02 \\%& commonly used \\ 
  Logic as Control Flow & 56.94\% & 0.95 \\%& commonly used \\ 
  Automatic Semicolon Insertion & 33.33\% & 0.13 \\%& commonly used \\ 
  Arithmetic as Logic & 23.61\% & 0.02 \\%& rarely used \\ 
  Comma Operator & 15.28\% & 0.03 \\%& rarely used \\
   \bottomrule
\end{tabular}
\label{tab:occurrences-summary}
\end{table}

@@

The results of our survey and interviews suggest that the Conditional Operator does not contribute significantly as a source of misunderstanding (increasing the number of wrong answers in 9\% of the cases, according to our survey). Nonetheless, the Post-Increment Expression and Automatic Semicolon Insertion atoms are listed in the top four sources of misunderstanding  (Table~\ref{tab:difference-correctness}), and also frequently appear in open source systems. As such, refactoring existing systems to avoid Post-Increment and Automatic Semicolon Insertion might improve the readabiliy of large amounts of code in JavaScript projects.

### 2021-10-22 main.tex
Made a pass through Section 4.

\begin{abstract}
  Evolving legacy code is a challenging task, particularly when the code has been poorly written or uses confuse idioms and language constructs, which might increase maintenance efforts and impose a significant cognitive load on developers. For this reason, researchers have investigated possible sources of confusion in codebases, including the impact of small code patterns (hereafter atoms of confusion) that contribute to misunderstanding the source code written in statically typed languages such as \clang, \cpplang, and Java. In this work, we investigate whether atoms of confusion identified in statically typed languages also confuse developers of a dynamically typed language. We use JavaScript as a representative example of dynamic programming languages and collect evidence from a mixed-method research effort: a survey, a set of interviews with practitioners, and a mining software repositories (MSR) study. Our survey and interviews confirm that atoms of confusion lead to code that is hard to understand in JavaScript. Considering ten atom candidates, developers correctly predict the outcome of at least 15\% more cases in code snippets where atom candidates are not present, compared to alternative versions that include the atom candidates. For five of these atoms, the difference is statistically significant (accounting for p-value correction). Effect size is large for four atoms and medium for one of them. In addition, our MSR effort reveals that atom candidates are frequent and used intensively in 72 popular open-source JavaScript systems. Four atom candidates appear in 90\% of the analyzed projects, and two of them occur more than once for every 100 lines of code in the dataset. Altogether, our findings might help practitioners: (1) better understand the implications of atoms of confusion on understanding JavaScript code, (2) avoid writing code that is unnecessarily difficult to maintain, and (3) design program transformation tools that remove potential sources of misunderstanding in JavaScript code.
\end{abstract}

>>>>

\begin{abstract}
  Evolving legacy code is a challenging task, particularly when the code has been poorly written or uses confuse idioms and language constructs, which might increase maintenance efforts and impose a significant cognitive load on developers. For this reason, researchers have investigated possible sources of confusion in codebases, including the impact of small code patterns (hereafter atoms of confusion) that contribute to misunderstanding the source code written in statically typed languages such as \clang, \cpplang, and Java. In this work, we investigate whether atoms of confusion identified in statically typed languages also confuse developers of a dynamically typed language. We use JavaScript as a representative example of dynamic programming languages and collect evidence from a mixed-method research effort: a survey, a set of interviews with practitioners, and a mining software repositories (MSR) study\castor{Do we still want to talk about the MSR part as a "study"? I remember Rodrigo saying that we did not.}. Our survey and interviews confirm that atoms of confusion lead to code that is hard to understand in JavaScript. Considering ten atom candidates, developers correctly predict the outcome of at least 15\% more cases in code snippets where atom candidates are not present, compared to alternative versions that include the atom candidates. For five of these atoms, the difference is statistically significant (accounting for p-value correction). Effect size is large for four atoms and medium for one of them. In addition, our MSR effort reveals that atom candidates are frequent and used intensively in 72 popular open-source JavaScript systems. Four atom candidates appear in 90\% of the analyzed projects, and two of them occur more than once for every 100 lines of code in the dataset. Altogether, our findings might help practitioners: (1) better understand the implications of atoms of confusion on understanding JavaScript code, (2) avoid writing code that is unnecessarily difficult to maintain, and (3) design program transformation tools that remove potential sources of misunderstanding in JavaScript code.
\end{abstract}
